Clustering of Distributions on 1-Dimensional Manifolds

### 2021-09-29 scratch/distributions-on-1d-manifolds.Rmd
added multidimensional case

\begin{theorem}
Let $X_1, ..., X_n \iid F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N \in \mathbb{N}$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{theorem}

@@

$$
\begin{aligned}
P(\max_i D_i < \delta) & = P(\text{all } D_i < \delta) \\
& = 1 - P(\text{some } D_i > \delta) \\
& \geq 1 - \sum_i^{n-1} P(D_i > \delta) \\
& = 1 + (n - 1) (1 - a \delta)^n - (n - 1) (1 - F(1 - \delta) - a \delta)^n
\end{aligned}
$$

This converges to $1$ in the limit $n \to \infty$.
\end{proof}

@@

\begin{theorem}
Let $X_1, ..., X_n \iid F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. Define $E_n$ as the event that every point on the support is within distance $\delta / 2$ to $X_i$ for $i \in [n]$. Then for any $\epsilon > 0$, there exists $N = O(-\log \epsilon)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
\end{theorem}

>>>>

\begin{theorem}
\ref{thm:generalized}
Let $X_1, ..., X_n \iid F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N \in \mathbb{N}$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{theorem}

@@

$$
\begin{aligned}
P(\max_i D_i < \delta) & = P(\text{all } D_i < \delta) \\
& = 1 - P(\text{some } D_i > \delta) \\
& \geq 1 - \sum_i^{n-1} P(D_i > \delta) \\
& = 1 - (n - 1) (1 - a \delta)^n + (n - 1) (1 - F(1 - \delta) - a \delta)^n
\end{aligned}
$$

This converges to $1$ in the limit $n \to \infty$.

We can also approximate $F(1 - \delta) \approx 1 - a \delta$, which yields 
$1 - (n - 1) (1 - a \delta)^n$. Setting this $\geq 1 - \epsilon$ ...
\end{proof}

@@

\begin{theorem}
\ref{thm:multidim}
Let $X_1, ..., X_n \iid F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
Define $E_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected. 
Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon \eta}{\log (1 - \frac{a \eta^r}{r^{r / 2}})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
\end{theorem}

\begin{proof}[Proof (sketch)]
Divide the hypercube $[0, 1]^r$ into a grid of sub-hypercubes of side length at most $\eta / \sqrt{r}$. 
$E_n$ is satisfied if each sub-hypercube contains at least one $X_i$ from the sample.

$$
\begin{aligned}
P(E_n) & = 1 - P(\text{some cells don't contain } X_i) \\
& \geq 1 - \sum_k^{\lceil \sqrt{r} / \eta \rceil} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
& \geq 1 - \lceil \sqrt{r} / \eta \rceil (1 - a \eta^r / r^{r/2})^n
\end{aligned}
$$

Setting this $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
\end{proof}

### 2021-09-29 scratch/distributions-on-1d-manifolds.tex
added multidimensional case

\begin{theorem}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N \in \mathbb{N}$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{theorem}

@@

\begin{theorem}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. Define $E_n$ as the event that every point on the support is within distance $\delta / 2$ to $X_i$ for $i \in [n]$. Then for any $\epsilon > 0$, there exists $N = O(-\log \epsilon)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
\end{theorem}

>>>>

\begin{theorem}
\ref{thm:generalized}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]$, and suppose $f(x)$ is continuous and $f(x) \geq a > 0$ everywhere on the support. 
Let $D_i = X_{(i+1)} - X_{(i)}$. 
Then for any $\epsilon > 0$, there exists $N \in \mathbb{N}$ such that $P(\max_i D_i < \delta) \geq 1 - \epsilon$ when $n > N$.
\end{theorem}

@@

\begin{theorem}
\ref{thm:multidim}
Let $X_1, ..., X_n \stackrel{\mathrm{iid}}{\sim}F$ with support $[0, 1]^r$, and $f(x) \geq a > 0$ everywhere on the support. 
Define $E_n$ as the event that an $\eta$-neighborhood graph constructed from the sample is connected. 
Then for any $\epsilon > 0$, there exists $N = O \bigg( \frac{\log \epsilon \eta}{\log (1 - \frac{a \eta^r}{r^{r / 2}})} \bigg)$ such that $P(E_n) > 1 - \epsilon$ when $n \geq N$.
\end{theorem}

\begin{proof}[Proof (sketch)]
Divide the hypercube $[0, 1]^r$ into a grid of sub-hypercubes of side length at most $\eta / \sqrt{r}$. 
$E_n$ is satisfied if each sub-hypercube contains at least one $X_i$ from the sample.

$$
\begin{aligned}
P(E_n) & = 1 - P(\text{some cells don't contain } X_i) \\
& \geq 1 - \sum_k^{\lceil \sqrt{r} / \eta \rceil} \prod_i^n P(X_i \text{ is not in the } k^{th} \text{ hypercube}) \\
& \geq 1 - \lceil \sqrt{r} / \eta \rceil (1 - a \eta^r / r^{r/2})^n
\end{aligned}
$$

Setting this $\geq 1 - \epsilon$ and solving for $n$ yields the desired rate. 
\end{proof}
